### 2017-09-25 draft1.1.tex
new project

%\documentclass[11pt]{article}
\documentclass[%
%reprint,
superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose,
preprint,
showpacs,
%preprintnumbers,
nofootinbib,
%nobibnotes,
%bibnotes,
amsmath,amssymb,
%aps,
%pra,
%prb,
prc,
%paper,
%rmp,
%prstab,
%prstper,
%twocolumn,
floatfix ]%
{revtex4-1}

\usepackage{comment}
%\usepackage[dvips]{graphicx}
%\usepackage{amsmath,amssymb}
%\usepackage{mathrsfs}
%\usepackage{txfonts}
\usepackage{braket}
%\usepackage{latexsym}
%\usepackage{ascmac}
%\usepackage{fancybox}
%\usepackage{type1cm}
\usepackage{cases}
\usepackage{caption}
\usepackage[top=25truemm,bottom=20truemm,left=20truemm,right=20truemm]{geometry}

  \makeatletter
    \renewcommand{\theequation}{%
    \thesection.\arabic{equation}}
    \@addtoreset{equation}{section}
  \makeatother

\begin{document}

\title{Comparative study of requantization of time-dependent mean-field
for pairing dynamics}
\author{Fang Ni}
 \affiliation{Faculty of Pure and Applied Sciences,
              University of Tsukuba, Tsukuba 305-8571, Japan}
\author{Takashi Nakatsukasa}
 \affiliation{Center for Computational Sciences,
              University of Tsukuba, Tsukuba 305-8577, Japan}
 \affiliation{Faculty of Pure and Applied Sciences,
              University of Tsukuba, Tsukuba 305-8571, Japan}
 \affiliation{iTHES Research Group, RIKEN, Wako 351-0198, Japan}

\begin{abstract}
In order to obtain quantal information, such as enegy eigenvalues
and transition matrix elements,
it is useful to requantize the mean-field dynamics.
We study the two-level pairing model to compare different
requantization methods and their accuracy.
Among the different methods, the one constructing microscopic
wave functions, using the TDHFB trajectories fulfilling the
Einstein-Brillouin-Keller quantization condition,
is the most accurate.
We also examine the performance of the collective model,
analogous to the Bohr model of the quadrupole vibrations and rotation,
assuming the pairing gap parameter as the collective coordinate.
It turns out that the applicability of such a model is very limited
for the pairing case.
\end{abstract}

\maketitle

\section{Introduction}

Pairing correlation is important 
in the ground state of open-shell nuclei.
Many evidences of the pairing correlation were observed in experiment,
including odd-even mass difference, moments of inertia of rotational bands,
and quasi-particle spectra in odd nuclei.
Even in closed-shell nuclei, the pairing dynamically plays an important
role in elementary modes of excitation, such as pairing vibrations
\cite{BM75, RS80, BB05}.
Properties of low-lying modes of excitation in even-even nuclei are
expected to be determined by interplay between
the pairing and the quadrupole correlations.
However, the true nature of the low-lying excitations is still unclear,
especilly for excited $J^\pi=0^+$ states \cite{garrett,HW11}.
Understanding the pairing dynamics is a key ingredient for solving
mysteries of excited $0^+$ states.

The ground states in many of even-even nuclei are well
described by the Bardeen-Cooper-Schriefer (BCS)
and the Hartree-Fock-Bogoliubov (HFB) theories \cite{RS80,BHR03}.
Its time-dependent version, 
the time-dependent HFB (TDHFB) theory \cite{BR86,NMMY16},
is a natural extension of the static HFB theory.
The small-amplitude limit of the TDHFB is known to be 
the quasiparticle random phase approximation (QRPA).
%Because of the improvement of computational power,
%dynamics and reaction process in nuclei have been investigated
%based on QRPA.
The QRPA has been extensively utilized and successfully
describe properties of giant resonances.
In contrast, many of low-lying excited states
cannot be well reproduced by QRPA \cite{NMMY16}.
This may be due to their slowly moving large-amplitude nature.

%Because the energy scale of pairing correlation is a few MeV, not only quadrupole deformation, but also pairing correlation play significant role in low-lying excited states. The importance of pairing in low-lying excited states is especially mentioned in rare earth region\cite{garrett}. The competition between quadrupole and pairing is considerable, hence the motion is expected to be large amplitude. \par
%The most conventional approach to describe large amplitude collective motion (LACM) in many-body system is time dependent Hartree-Fock(-Bogoliubov) TDHF(B) theory.

The Bohr Hamiltonian was introduced to describe low-energy
nuclear collective motion in quadrupole degrees of freedom with
the deformation parameters $(\beta, \gamma)$ and the Euler angles
$(\alpha,\beta,\gamma)$.
For the pairing motion, the collective coordinates are analogously
chosen as the pair deformation (gap) $\Delta$ and the gauge angle $\Phi$
\cite{BBPK70}.
Based on the pairing collective Hamiltonian, 
effects of the pair motion on the quadrupole vibrations %and fission half lives
have been discussed in literatures \cite{delta1,delta3}.
%\cite{delta1,delta2,delta3}. 
However, it is not trivial whether the pair deformation is really a suitable
choice for the collective coordinate,
which should be investigated in a microscopic approach based on
the TDHFB dynamics.

In principle, the TDHFB dynamics can be large amplitude motion.
The problem is that it is not easy to determine quantum mechanical 
quantities, such as energy eigenvalues and transition matrix elements,
from the TDHFB trajectories.
For this purpose,
since the TDHF(B) trajectory corresponds to a classical limit of the path
integral formulation,
the requantization of the mean-field dynamics was
proposed \cite{Neg80,path,path2,Rei80}.
It recovers quantum fluctuations missing in the mean-field
level, and possibly enables us to describe large-amplitude collective
tunneling phenomena such as spontaneous fission \cite{Neg80}.
The requantization of the TDHFB is particularly feasible
for integrable systems, because the system is described by separable
action-angle variables $(I_k, \phi_k)$, leading to
the Einstein-Brillouin-Keller (EBK) quantization condition.
However, for nonintegrable systems in general, it is difficult to find
suitable periodic orbits to quantize.
A possible solution to this difficulty is to find a decoupled collective
subspace spanned by
a single coordinate and its conjugate momentum \cite{NMMY16}.
Since the one-dimensional system is integrable, the quantization is
practicable.

Our final goal is to study the role of dynamical pairing effects
and to reveal nature of the mysterious excited $0^+$ states.
As the first step Toward this goal, we investigate accuracy and
applicability of the requantized TDHFB model for a two-level pairing model
with equal degeneracy $\Omega$ \cite{HF60},
especially on the calculation of two-particle-transfer matrix elements.
In Ref.~\cite{CDS84},
the two-particle-transfer matrix elements were eveluated as
Fourier components of the time-dependent mean values of the pair
creation operators,
which demonstrates nice agreement with the exact results at
large $\Omega$ values ($\Omega=40$).
However, in the realistic values of $\Omega$, we will show that
the deviation is prominent.
The collective model of the pairing motion, which assumes the pair deformation
as the collective coordinate, has a similar tendency, namely, 
applicability limited to large $\Omega$ cases \cite{BBPK70}.
This deficiency is mainly due to the small collectivity in the pairing motion
in realistic situations.
In this paper, in order to improve the quantitative estimate of the matrix
elements, we construct microscopic wave functions
based on the EBK quantization for the integrable systems.
The wave functions are obtained
from the stationary phase approximation for the path-integral form \cite{SM88}.
Its superiority to the other methods
becomes more evident for smaller values of $\Omega$.

%Nuclear collective model has the advantage to deal with the requantization. The collective coordinates can be easily quantized. Also, the mass parameters and collective potential with respect to the collective coordinates can be derived from microscopic theory (e.g. constrained HFB). For example, Bohr Hamiltonian, which can describe nuclear collective motion in quadrupole degrees of freedom, is written by collective coordinates of quadrupole deformation parameter $\beta, \gamma$ and Euler angle $\theta_i \ (i=1,2,3)$. For pairing LACM (PLACM), the collective coordinates are gauge angle $\Phi$ and pairing gap $\Delta$. $\Delta$ corresponds to deformation and $\Phi$ corresponds to rotation in U(1) gauge space. The concept of pairing collective coordinates is firstly introduced in Ref. \cite{first}. Based on pairing collective Hamiltonian, the pairing degrees of freedom influence not only the value of pairing gap in ground state, but also transition probability and spontaneous fission half-lives\cite{delta1,delta2,delta3}. On the other hand, without pairing collective Hamiltonian, there is little study about PLACM in requantization of TDHFB. Because the pairing collective coordinates are chosen ``by hand'', we need to investigate the validity of such collective coordinates.\par

%The main purpose of this paper is elucidate PLACM in two-level system of a simple nuclear pairing model -- Richardson model.  We consider all of the canonical coordinates in TDHFB trajectory. Such canonical coordinates can be generated from Thouless' theorem. In realistic nuclear system, we have huge number of canonical coordinates. However, the number of canonical coordinates is limited in Richardson model, which has advantage to be studied for the requantization of TDHFB without collective model. Especially, two-level system contains the least degrees of freedom for discussing about pairing rotation and pairing vibration. We also transform the canonical coordinates to $\Delta$ and $\Phi$, to discuss whether the pairing collecitve Hamiltonian is validity. \par
%To describe LACM from microscopic picture, the major approach is constrained Hartree-Fock(-Bogoliubov) (CHFB) or constrained BCS (CBCS) method by assuming collective coordinates as the constraint condition. \par
%Because the collective of gauge angle $\Phi$ and pairing gap $\Delta$ is chosen ``by hand'', it is suspective whether the system prefers the motion along artificial collective coordinates. We need to consider some more ``natural'' collective coordinates.\par

The paper is organized as follows. 
In Sec. \ref{sec:TDHFB},
we derive a TDHFB classical Hamiltonian for the pairing model.
In Sec. \ref{sec:requantization}, 
the requantization of the TDHFB is performed using different methods,
based on the canonical and the EBK quantizations.
In Sec. \ref{sec:results}, 
the numerical results for the two-level model are shown and compared with
exact results.
Properties of the pairing collective coordinate is also discussed.
We give a conclusion in Sec. \ref{sec:conclusion}.

%\section{Classical picture}
\section{Classical form of TDHFB Hamiltonian}
\label{sec:TDHFB}

The Hamiltonian of the pairing model is given in terms of
single-particle energies $\epsilon_l$ and the pairing strength $g$ as
\begin{equation}
%  H &= \sum_l \epsilon_l n_l - \frac{g}{4} \sum_{ll'} P_l^{\dag}P_{l'}
	H = \sum_l \epsilon_l n_l - g \sum_{l,l'} S_l^+ S_{l'}^-
    = \sum_l\epsilon_l(2S_l^0+\Omega_l) - g S^+ S^{-} ,
\end{equation}
where we use the SU(2) quasispin operators,
$\boldsymbol{S}=\sum_l \boldsymbol{S}_l$, with
\begin{eqnarray}
        S_l^0 &=& \frac{1}{2}(\sum_ma_{lm}^{\dag}a_{lm}-\Omega_l) ,\\
        S_l^{+} &=& \sum_{m>0}a_{lm}^{\dag}a_{l\overline{m}}^{\dag} ,
\quad   S_l^{-} = S_l^{+\dag} .
\end{eqnarray}
Each single-particle energy $\epsilon_l$ possesses $(2\Omega_l)$-fold
degeneracy ($\Omega_l=j_l+1/2$)
and $\sum_{m>0}$ indicates the summation over $m=1/2,3/2,\cdots,\Omega_l-1/2$.
The occupation number of each level $l$ is given by
$
	n_l = \sum_m a^{\dag}_{lm}a_{lm} = 2S_l^0+\Omega_l ,
$.
%The first term giving a spherical single particle energy $\epsilon_l$ for each j-shell, which 
The quasispin operators satisfy the commutation relations
\begin{equation}
  [S_l^0,S_{l'}^{\pm}] = \pm\delta_{ll'}S_{l}^{\pm},
	\quad [S_{l}^{+},S_{l'}^{-}] = 2\delta_{ll'}S_{l}^{0} .
\end{equation}
The magnitude of quasispin for each level is
$S_l=\frac{1}{2}(\Omega_l-\nu_l)$, where $\nu_l$ is the seniority
quantum number, namely the number of unpaired particle at each level $l$.
%We call $\nu=\sum_l \nu_l=0$ excited states as pairing collective excited states. The pairing collective excited states can be simple $2p2h, 4p4h\cdots$ excitation or the complex combination of $2p2h, 4p4h\cdots$ components.
In the present study, we only consider seniority zero states with
$\nu=\sum_l \nu_l=0$.
The residual two-body interaction only consists of monopole pairing
interaction which couples two particles to zero angular momentum.
We obtain exact solutions either by solving Richardson equation
\cite{Richardson,Richardson2,Richardson3} or
by diagonalizing the Hamiltonian using the quasispin symmetry.

%\subsection{Coherent state}
\subsection{Coherent-state representation of the TDHFB Hamiltonian}

The coherent state for the seniority $\nu=0$ states
($S_l=\Omega_l/2$) is constructed as
\begin{equation}
	\ket{Z(t)} = \prod_{l} \left(1+|Z_l(t)|^2\right)^{-\Omega_l/2}
	\exp (Z_l(t) S_l^{+}) \ket{0}
 \label{coherent}
\end{equation}
where $\ket{0}$ is the vacuum (zero particle) state,
$Z_l(t)$ are time-dependent complex variables which describe
motion of the system. 
In the SU(2) quasispin representation,
$\ket{0}=\prod_l \ket{S_l,-S_l}$.
The coherent state $\ket{Z(t)}$ is a superposition of
states with different particle numbers
without unpaired particles.
%If $Z_l$ are all real,
In the present pairing model,
the coherent state is the same as the time-dependent BCS wave function
with $Z_l(t)=v_l(t)/u_l(t)$,
where $(u_l(t),v_l(t))$ are the time-dependent BCS $u,v$ factors.

The TDHFB equation can be derived from the tiem-dependent variational
principle,
\begin{equation}
	\delta \mathcal{S} = 0,
	\quad
%  \delta \int \braket{\phi(t)|i\frac{\partial}{\partial t}-H|\phi(t)}dt = 0 ,
	\mathcal{S}\equiv
	\int \braket{\phi(t)|i\frac{\partial}{\partial t}-H|\phi(t)}dt,
  \label{TDHFB}
\end{equation}
where $\ket{\phi(t)}$ is the time-dependent generalized Slater determinant.
In the present case, we adopt the coherent state of Eq. (\ref{coherent}),
$\ket{\phi(t)}=\ket{Z(t)}$.
%Keeping the time-dependent state vector fulfills normalization $\braket{Z|Z}=1$, 
The action $\mathcal{S}$ is
\begin{equation}
	\mathcal{S}  = \int \mathcal{L}(t) dt
	%&= \int \braket{\phi(t)|i\frac{\partial}{\partial t}-H|\phi(t)}dt \\
	= \int dt \left\{ \frac{i}{2} \sum_l \frac{\Omega_l}{1+|Z_l|^2}
  (Z_l^*\dot{Z_l}-Z_l\dot{Z_l^*}) - \braket{Z|H|Z} \right\} ,
\label{S}
\end{equation}
and
\begin{align}
%  \braket{Z|H|Z} &= \sum_l \epsilon_l \left\{ 2S_l\left( 1-\frac{2}{1+|Z_l|^2} \right) + \Omega_l \right\}
%  -g \sum_l \frac{2S_l|Z_l|^2(2S_l+|Z_l|^2)}{(1+|Z_l|^2)^2}
%  -g \sum_{l_1 \neq l_2} \frac{2S_{l_1}Z_{l_1}}{1+|Z_{l_1}|^2}\frac{2S_{l_2}Z_{l_2}^*}{1+|Z_{l_2}|^2}.
  \braket{Z|H|Z} &= \sum_l \epsilon_l \frac{2\Omega_l |Z_l|^2}{1+|Z_l|^2}
  -g \sum_l \frac{\Omega_l|Z_l|^2(\Omega_l+|Z_l|^2)}{(1+|Z_l|^2)^2}
  -g \sum_{l_1 \neq l_2} \frac{\Omega_{l_1}Z_{l_1}}{1+|Z_{l_1}|^2}\frac{\Omega_{l_2}Z_{l_2}^*}{1+|Z_{l_2}|^2}.
	\label{TDHFB_Hamiltonian_1}
\end{align}
We transform the complex variables $Z_l$ into real variables $(q_l,\chi_l)$
by $Z_l = \tan{\frac{\theta_l}{2}}e^{-i\chi_l}$ and $q_l=\cos{\theta_l}$.
The Lagrangian $\mathcal{L}$ and the expectation value of Hamiltonian become
\begin{align}
\mathcal{L}(t) &= \sum_l \frac{\Omega_l}{2}
	(1-q_l)\dot{\chi_l} - \mathcal{H}(Z,Z^*) ,\\
	\mathcal{H}(Z,Z^*)&\equiv \braket{Z|H|Z} = \sum_l \epsilon_l\Omega_l(1- q_l) - \frac{g}{4}\sum_l \Omega_l (\Omega_l(1-q_l^2)+(1-q_l)^2) \nonumber \\
&- \frac{g}{4}\sum_{l_1\neq l_2} \Omega_{l_1}\Omega_{l_2}\sqrt{(1-q_{l_1}^2)(1-q_{l_2}^2)}e^{-i(\chi_{l_1}-\chi_{l_2})}   .
\label{TDHFB_Hamiltonian_2}
\end{align}
Here, $\chi_l$ represent the gauge angle of each level,
and $q_l$ are related to the occupation probability, $q_l=|u_l|^2-|v_l|^2$.
Since the Hamiltonian (\ref{TDHFB_Hamiltonian_2}) depends only on
relative difference in the gauge angles,
the ``global'' gauge angles, $\Phi\propto\sum_l \chi_l$,
is a cyclic variable.

\subsection{Two-level case}

In a two-level system,
it is convenient to define global and relative gauge angles,
$\Phi$ and $\phi$, respectively.
\begin{align}
  \Phi &\equiv \frac{\chi_1 + \chi_2}{2}, \quad\quad
\phi\equiv \chi_2 - \chi_1,
\label{phi}
\end{align}
whose ranges are $0\leq \Phi \leq 2\pi$ and $-2\pi \leq \phi \leq 2\pi$.
Their conjugate momenta $(J,j)$ are given by
\begin{align}
	J &= \frac{\partial\mathcal{L}}{\partial\dot{\Phi}} =  \sum_{l=1}^2 \frac{\Omega_l}{2}(1-q_l), 
\quad\quad
j = \frac{\partial\mathcal{L}}{\partial\dot{\phi}} = \frac{\Omega_2(1-q_2) - \Omega_1(1-q_1)}{4}. \label{pi} .
\end{align}
Calculating the occupation number $n_l$ in the level $l$,
the physical meaning of these conjugate momenta becomes obvious.
\begin{equation}
   n_l = \braket{Z|n_l|Z} % = -2S_lq_l+\Omega_l \nonumber  \\
	= \Omega_l(1-q_l) .
\end{equation}
Therefore, $J$ corresponds to the total particle number
$N=\sum_l n_l$,
while $j$ corresponds to the difference of occupation number
between upper level and lower level. 
\begin{align}
  J &= \frac{N}{2}, \hspace{5mm} j = \frac{n_{2}-n_{1}}{4} .
  \label{momentum}
\end{align}
Since the coherent states (\ref{coherent}) contains only even number of
particles in each level,
we assume that $n_l$ are even numbers.
%The expectation value of total particle number is
%\begin{align}
%   N &= \braket{Z|N|Z} = \sum_l (-2S_lq_l+\Omega_l) \nonumber  \\
%	&= 2\sum_l S_l(1-q_l)+\nu .
%\end{align}
%Also, the expectation value of occupation number in each level is
The Hamiltonian in terms of these canonical variables
$(\phi,j;\Phi,J)$ is given by
\begin{equation}
\mathcal{H}(\phi,j;J)
	= \sum_{l=1,2} \Omega_l \left\{ \epsilon_l(1- q_l)
	- \frac{g}{4} (\Omega_l(1-q_l^2)+(1-q_l)^2) \right\}
- \frac{g}{2}\Omega_1\Omega_2
	\sqrt{(1-q_1^2)(1-q_2^2)} \cos\phi .
\label{TDHFB_Hamiltonian_3}
\end{equation}
with
\begin{equation}
	q_l=\frac{\Omega_l - J -2(-1)^l j}{\Omega_l} 
	\quad \mbox{ for } l=1,2.
	\label{q_l}
\end{equation}
Note that the Hamiltonian does not depend on the global gauge angle $\Phi$.
This leads to the particle number conservation, $dN/dt=0$.

%\begin{align}
%  \mathcal{H}(\phi(t),\pi(t);\Phi(t),\Pi(t)) =& \braket{Z|H|Z} \nonumber \\
%  =& \sum_{l=1,2} \epsilon_l(-2S_lq_l+\Omega_l) - g\sum_{l=1,2} \frac{S_l}{2}(2S_l(1-q_l^2)+(1-q_l)^2) \nonumber \\
%  &- 2g S_{1}S_{2}\sqrt{(1-q_{1}^2)(1-q_{2}^2)}\cos{\phi}.
%\label{hamiltonian}  
%\end{align}
%In (\ref{hamiltonian}), there is no $\Phi$ dependence in Hamiltonian, hence the expectation value of total particle number is conserved. The dependence of classical Hamiltonian becomes
%\begin{align}
%  \mathcal{H}(\phi(t)\},\pi(t);\Phi(t),\Pi(t))=\mathcal{H}(\phi(t),\pi(t),\Pi) .
%\end{align}
%\par

%The motion on $\Phi$ direction corresponds to pairing rotation. In quantum state, pairing rotational mode is labeled by quantum number $N$. The excitation states emerges in $N+2, N-2,\cdots$ system. In single-j system, the energy spectra is completely parabolic with respect to $N$. However, it contains higher order of $N$ from (\ref{hamiltonian}) in general system. \par
%Excluding the cyclic variable, all we have to study is the one dimensional motion on $\phi$ direction, which corresponds to pairing vibration. The excitation states in pairing vibrational mode emerges in the same $N$ system. We can understand the classical picture of pairing vibration in two dimensional phase space. The classical trajectory fulfills

The TDHFB equation can be written in a form of
the classical equations of motion:
\begin{eqnarray}
	&&\frac{d\Phi}{dt} = \frac{\partial\mathcal{H}}{\partial J} ,\quad\quad
\frac{dJ}{dt} = -\frac{\partial\mathcal{H}}{\partial \Phi} , \\
	&&\frac{d\phi}{dt} = \frac{\partial\mathcal{H}}{\partial j} ,\quad\quad
\frac{dj}{dt} = -\frac{\partial\mathcal{H}}{\partial \phi} .
\label{TDHFB_equation}
\end{eqnarray}
Since the $J=N/2$ and the total energy $E$ are constants of motion,
the TDHFB trajectories with given $N$ and $E$ are determined
in the two-dimensional phase space $(\phi,j)$ with the condition
\begin{equation}
  \mathcal{H}(\phi(t),j(t);J=N/2) = E.
\end{equation}
Examples of the classical trajectories in the phase space
$(\phi,j)$ are shown in Fig.~\ref{fig:phase_space}.
The figure shows contour lines of energy for systems,
which correspond to the TDHFB trajectories,
with $\Omega_1=\Omega_2=8$ and $N=16$ with different values of $g$.
%$N=16$ system with energy contour lines fulfilling (\ref{Sommerfeld}). 
The transition from normal to superfluid phase takes place at
$g=\Delta\epsilon(2\Omega)^{-1}$ with $\Delta\epsilon=\epsilon_2-\epsilon_1$.
Using a dimensionless parameter
$
  x = 2g\Omega/\Delta\epsilon,
$
at $x<1$, the ground state is normal with fully occupied
lower level $n_1=N$ and empty upper level ($n_2=0$).
All the TDHFB trajectories represent rotational behavior with respect to
relative angle $\phi$.
At $g=x=0$, the Hamiltonian becomes independent from $(\Phi,\phi)$, then,
the occupation numbers, $n_1$ and $n_2$, are constants of motion.
At $x>1$, the energy-minimum point and the closed trajectories appear around
%$-N/4 < \pi < N/4$ and $\phi=0$,
$j=j_0$ ($-N/4 < j_0 < N/4$) and $\phi=0$,
which suggests the vibrational behavior for $\phi$.
At higher energies, the trajectories become open (rotational-like),
which suggests a phase transition from super to normal phases
as a function of excitation energy.
%We have two types of trajectory, opened trajectory and closed trajectory. With BCS ground state, the closed trajectory which corresponds to a kind of condensed state may emerge in low-lying region. There is a phase transition point from closed trajectory to opened trajectory. Near the phase transition point, the fluctuation is important to deal with quantum states. The fluctuation causes the difficulty of the description by TDHFB theory.

In the single-$j$ model, since the second term in the Hamiltonian
(\ref{TDHFB_Hamiltonian_3}) is absent,
the Hamiltonian is exactly quadratic with respect to $N$,
$\mathcal{H}(N) = \mathcal{H}_0+(N-N_0)^2/(2\cal{J})$.
The moment of inertia for the pair rotation is 
$\mathcal{J}=2/g\times\Omega/(\Omega-1)\approx 2/g$
at $\Omega\rightarrow\infty$.
In the multi-$j$ case, the Hamiltonian contains higher-order
terms in general.

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=60mm,angle=-90]{images/phase.eps}
 \end{center}
 \caption{Energy contour plot for $\Omega_1=\Omega_2=8$ and $N=16$.
	The lines indicate the TDHFB trajectories fulfilling
	the EBK quantization condition of Eq. (\ref{EBK}).
	}
 \label{fig:phase_space}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Requantization of TDHFB for the two-level pair model}
\label{sec:requantization}

In order to determine energy eigenstates,
we need to requantize the TDHFB trajectories.
%The straightforward method is canonical quantization.  However, the canonical quantization has problem to deal with two-particle transition (see section 4).  To solve the problem, we also consider semiclassical approach, which is the approximation of path integral.
Since the present two-level model is integrable (Sec.~\ref{sec:TDHFB}),
it is feasible to apply the stationary-phase approximation to
the path integral expression.
In addition, we also study the canonical quantization of the TDHFB
Hamiltonian, and the matrix elements extracted by the Fourier components
\cite{CDS84}.
We show results of these different approaches to the pairing model.

\subsection{Stationary phase approximation to the path integral}
\label{sec:SPA}

Starting an arbitrary state $\ket{\psi(0)}$ at time $t=0$,
the time-dependent full quantum state can be written in the path integral form
\begin{equation}
\ket{\psi(t)}
	= e^{-iHt} \ket{\psi(0)}
	= \int d\mu(Z'') \ket{Z''} \int d\mu(Z')
	\int_{Z(0)=Z'}^{Z(t)=Z''} \mathcal{D}\mu[Z(\tau)])
	e^{i\mathcal{S}[Z(\tau)]} \psi(Z') ,
\label{path_integral_expression}
\end{equation}
where $\psi(Z)\equiv \braket{Z|\psi(0)}$ and
the invariant measure $d\mu(Z)$ is defiend by 
the unity condition,
\begin{equation}
  \int d\mu(Z) \ket{Z}\bra{Z} = 1 .
  \label{unity}
\end{equation}
In Eq. (\ref{path_integral_expression}),
$\mathcal{S}[Z(\tau)]$ is the action (\ref{S}) along a given path $Z(\tau)$
with the initial coherent state $\ket{Z(0)}=\ket{Z'}$ and
the final state $\ket{Z(t)}=\ket{Z''}$,
then, the integration $\int \mathcal {D}\mu[Z(\tau)]$
is performed over all possible paths $\ket{Z(\tau)}$ between them.
Among all trajectories in the path integral,
the lowest stationary phase approximation selects
the TDHFB (classical) trajectories.
%\begin{align}
%	\int \mathcal{D}\mu(Z) e^{i\mathcal{S}(Z,Z_0)} \approx e^{i\mathcal{S}_{\rm cl}(Z,Z_0)} .
%\end{align}
\begin{equation}
	\ket{\psi(t)} \approx \int d\mu(Z') \ket{Z'_{\rm cl}(t)}
	e^{i\mathcal{S}_{\rm cl}(Z'_{\rm cl}(t),Z')} \psi(Z') ,
	\label{SPA}
\end{equation}
where the TDHFB trajectory starting from $\ket{Z'}$
ends at $\ket{Z'_{\rm cl}(t)}$ at time $t$.
The action $\mathcal{S}_{\rm cl}(Z_f,Z_i)$ is calculated along this
classical trajectory
connecting $Z_i=Z'_{\rm cl}(0)=Z'$ and $Z_f=Z'_{\rm cl}(t)$.
\begin{equation}
	\mathcal{S}_{\rm cl}(Z'_{\rm cl}(t),Z') \equiv \int_{0}^{t}
\braket{Z_{\rm cl}(t)|
	i\hbar\frac{\partial}{\partial t} - H
	|Z_{\rm cl}(t)} dt
	= \mathcal{T}[Z_{\rm cl}] - \mathcal{H}(Z',Z'^*) t 
	,
	\label{S_cl}
\end{equation}
with
\begin{equation}
\mathcal{T}[Z_{\rm cl}]\equiv
\int_{0}^{t} \braket{Z_{\rm cl}(t)| i\hbar\frac{\partial}{\partial t}
	|Z_{\rm cl}(t)} dt 
	= \int_{Z'}^{Z'_{\rm cl}(t)}
	\frac{i}{2} \sum_l \frac{\Omega_l}{1+|Z_l|^2}
  (Z_l^* dZ_l - Z_l dZ_l^*) .
\end{equation}
In the last equation of Eq. (\ref{S_cl}),
we used the fact that the TDHFB trajectory conserves the energy,
$\mathcal{H}(Z_{\rm cl}(t),Z_{\rm cl}^*(t))= \mathcal{H}(Z',Z'^*)$.
%We assume that only classical (TDHFB) trajectory contributes to the time dependent wave function

The energy eigenstates correspond to stationary states,
$\braket{Z|\psi(t)} \propto \braket{Z|\psi(0)}=\psi(Z)$,
which can be constructed by superposing the coherent states along
a periodic TDHFB trajectory $Z_{\rm cl}^{(k)}$ as \cite{path,path2,SM88}.
\begin{equation}
	\ket{\phi_k} = \oint d\mu(Z_{\rm cl}^{(k)}) \ket{Z_{\rm cl}^{(k)}}
	e^{i\mathcal{T}[Z_{\rm cl}^{(k)}]} .
\end{equation}
The single-valuedness of the wave function leads to
the quantization condition ($k$: integer)
\begin{equation}
	\mathcal{T}_\circ[Z_{\rm cl}^{(k)}]=
	\oint \frac{i}{2} \sum_l \frac{\Omega_l}{1+|Z_l^{(k)}|^2}
	\left(Z_l^{(k)*} dZ_l^{(k)} - Z_l^{(k)} dZ_l^{(k)*}\right) = 2k\pi .
	\label{EBK}
\end{equation}
The state evolves in time as $\ket{\phi(t)}=\ket{\phi}e^{-iE_k t}$,
with the energy of the $k$-th periodic trajectory,
$E_k=\mathcal{H}(Z_{\rm cl}^{(k)},Z_{\rm cl}^{(k)*})$.

Finding TDHFB trajectories satisfying the quantization condition
(\ref{EBK}) is an extremenly difficult task in general.
It is better founded and more practical
if the classical system is completely integrable.
In integrable systems, $M$ complex variables $Z(t)$ can be transformed into
the action-angle variables;
\begin{equation}
	Z(t) = \{Z_l(t);l=1,\cdots,M \} \rightarrow
	\{E; v_1,\cdots,v_{M-1}; \theta_1(t),\cdots,\theta_M(t)\} ,
\end{equation}
where the variables $E$ and $v$ define an invariant torus,
while $\theta(t)$ parameterize the coordinates on the torus.
The integration path of Eq. (\ref{EBK}) is now taken as topologically
independent closed path on the torus,
namely the EBK quantization condition.
There are $M$ independent closed paths and $M$ quantum numbers,
$k=(k_1,\cdots,k_M)$, to specify the stationary energy eigenstate.
These are associated with $M$ invariant variables,
$\{E_k;v_1^{(k)},\cdots,v_{M-1}^{(k)}\}$.
Using the invariant measure
\begin{align}
%  d\mu(Z) = \rho(Z)dEdv_1\cdots dv_{M-1}d\theta_1\cdots d\theta_M,
  d\mu(Z) = \rho(E,v,\theta) dEdv_1\cdots dv_{M-1}d\theta_1\cdots d\theta_M,
\end{align}
the $k$-th semiclassical wave function can be calculated as
\begin{equation}
  \ket{\phi_k} \propto \oint d\theta_1\cdots \oint d\theta_M \ 
	\rho(E_k,v^{(k)},\theta)
	\ket{E_k,v^{(k)},\theta} e^{i\mathcal{T}[E_k,v^{(k)},\theta]/\hbar} .
  \label{semi_wave_func}
\end{equation}
Here, we omit the integration with respect to the invariant variables,
$E$ and $v$.

%The wave function can be obtained after summing over whole phase space. In integrable system, we can write down the coordinates in $2M$ dimensional phase space as
%\begin{align}
%  Z = \{Z_{\mu}\} = \{E; v_1,\cdots,v_{M-1}; \theta_1(t),\cdots,\theta_M(t)\}
%\end{align}
%where $E$ is total energy, $v_i$ are invariant physical numbers and $\theta_i(t)$ are coordinates in torus. Having coordinate $q$ and conjugate momentum $p$, the Einstein-Brillouin-Keller (EBK) quantization condition is
%\begin{align}
%  \oint_{H(p,q)=E} \sum_i p_i dq_i = (n+\frac{\nu}{4})h  \hspace{2em} (n=1,2,\cdots)
%\end{align}
%where $h$ is plank constant and $\nu$ is the number of turning points in the trajectory. The EBK quantization condition attributes to Sommerfeld quantization
%\begin{equation}
%  \oint_{H(p,q)=E} p dq = (n+\frac{\nu}{4})h \hspace{2em} (n=1,2,\cdots),
%  \label{Sommerfeld}
%\end{equation}
%in one dimensional dynamics.

We apply semiclassical approach to the two-level pairing model.
The invariant measure in SU(2)$\otimes$SU(2) is 
\begin{align}
%	d\mu(Z) &= \prod_{l=1,2} \frac{S_l+1}{\pi}(1+|Z_l|^2)^{-2}d{\rm Re}Z_ld{\rm Im}Z_l \\
%  &= \prod_l \frac{S_l+1}{4\pi}\sin{\theta_l}d\theta_ld\phi_l \\
%  &= \frac{S_1+1}{4\pi}\frac{S_2+1}{4\pi}\frac{1}{S_1S_2}d\Phi d\Pi d\phi d\pi .
d\mu(Z) &= \prod_l \frac{\Omega_l+2}{2\pi}(1+|Z_l|^2)^{-2}d{\rm Re}Z_ld{\rm Im}Z_l \\
  &= \prod_l \frac{\Omega_l+2}{8\pi}\sin{\theta_l}d\theta_ld\phi_l \\
	&= \left(\prod_{l=1,2} \frac{1+2\Omega_l^{-1}}{4\pi}\right)
             d\Phi dJ d\phi dj .
\end{align}
In the last equation, we transform the canonical coordinates
by Eqs. (\ref{phi}) and (\ref{pi}).
%In two-level system, we can transform the canonical coordinates by (\ref{phi}) and (\ref{pi}).
%After coordinate transformation, the Hamilton's equations are
%\begin{align}
%\left\{
%\begin{array}{c}
%  \dot{\Phi} = \frac{\partial\mathcal{H}}{\partial \Pi} \\
%  \dot{\Pi} = -\frac{\partial\mathcal{H}}{\partial \Phi} 
%\end{array}
%\right. \hspace{10mm} \left\{
%\begin{array}{c}
%  \dot{\phi} = \frac{\partial\mathcal{H}}{\partial \pi} \\
%  \dot{\pi} = -\frac{\partial\mathcal{H}}{\partial \phi} 
%\end{array}
%\right. , \label{Hamilton}
%\end{align}
%and the measure and action integral become
%\begin{align}
%  d\mu(Z) &= \frac{S_1+1}{4\pi}\frac{S_2+1}{4\pi}\frac{1}{S_1S_2}d\Phi d\Pi d\phi d\pi \\
%  \mathcal{T}(Z) &=  \int (\Pi d\Phi+\pi d\phi) = \Pi\Phi + \int \pi d\phi.
%\end{align}
%The above canonical variables contain an invariant number $\Pi$. Because two-level system has 2*2 dimensional phase space, we need to find another conservative number before applying (\ref{semi_wave_func}). This task can be accomplished by canonical transformation 
Since the particle number $J=N/2$ and the total energy $E$ are invariant,
the two-level pairing model is integrable.
Thus, we can construct the semiclassical wave function using
Eq. (\ref{semi_wave_func}).
%This task can be accomplished by canonical transformation 
%\begin{align}
%  (\phi, \pi)  \to (t, E)
%\end{align}
%where $E$ is total energy in classical system.
The action integral is given by
\begin{equation}
	\mathcal{T}_k(\Phi,\phi;J)
%	= \Pi\Phi + \int \pi(t) \frac{d\phi}{dt} dt .
	= J\Phi + \int^\phi_{-\pi} j' d\phi' 
	= \frac{N}{2}\Phi + \int^t_0   %_{\mathcal{H}(\phi,\pi)=E_k}
	j(t') \frac{d\phi}{dt'} dt' 
\equiv \mathcal{T}_{N,E_k}(\Phi,t) ,
\end{equation}
where the integration $\int j d\phi$ is performed on the $k$-th
closed trajectory of Eq. (\ref{EBK}), and
the variables $(\phi,j)$ are transformed into $(t,E)$.
%where $\pi(t)$ can be solved from the right side of (\ref{TDHFB_equation}). (i.e. 2D Runge-Kutta method) \par
The semiclassical wave function fulfilling the EBK quantization condition
becomes
\begin{align}
	\ket{\phi_k^N} &\propto \oint d\Phi \oint dt
	e^{i\mathcal{T}_{N,E_k}(\Phi, t)}
	\ket{\Phi,t}_{N,E_k} \nonumber \\
%  &\propto \sum_k \frac{1}{k!}\int_0^{2\pi}d\Phi e^{i(\Pi-k)\Phi}\sum_{j=0}^{k}C_j^{(E,k)}\ket{S_1,-S_1+j;S_2,-S_2+(k-j)} \label{semi_wave_func1}
  &\propto \sum_{m=0}^{J}C_m^{(E_k,J)}\ket{S_1,-S_1+m;S_2,-S_2+(J-m)} ,
	\label{semi_wave_func1}
\end{align}
with $S_l=\Omega_l/2$, $J=N/2$, and the coefficients
\begin{align}
  C_m^{(E_k,J)} &= {}_J C_m\int_0^T dt \exp{\left( i\int^t j(t') \dot{\phi}(t') dt'-i(J/2-m)\phi(t) \right)}A(q_1,S_1,m)A(q_2,S_2,J-m) ,\\
  & \hspace{5mm} A(q,S,m) = \left(\frac{1-q}{2}\right)^{m/2}\left(\frac{1+q}{2}\right)^{S-m/2}\sqrt{\frac{(2S)!m!}{(2S-m)!}}, \hspace{5mm} \nonumber
\end{align}
where $T$ is the period of the closed trajectory.
The TDHFB-requantized wave functions (\ref{semi_wave_func1}) are
eigenstates of the total particle number.
This is due to the integration over the global gauge angle $\Phi$,
which makes the particle number projection not only for the ground state
but also for excited states.
See Appendix for detailed derivation of Eq.~(\ref{semi_wave_func1}).

Since we obtain the microscopic wave function of every eigenstate,
the expectation values and the transition matrix elements for any operator
can be calculated in a straightforward manner.
In Sec.~\ref{sec:results}, we show those of the pair-addition operator
$S^+$ which characterize properties of the pair condensates.

% In our system, we have two types of trajectory (Figure \ref{trajectory}). For opened trajectory, we define the size which is surrounded by one periodic trajectory and horizontal axis. Under this definition, the size changes smoothly when the trajectory changes from closed to opened. 
%
%\begin{figure}[htbp]
% \begin{minipage}{0.5\hsize}
% \begin{center}
%%  \includegraphics[bb=0 0 120 120]{images/Sommerfeld1.png}
% \end{center}
% \end{minipage}
% \begin{minipage}{0.5\hsize}
% \begin{center}
%%  \includegraphics[bb=0 0 120 120]{images/Sommerfeld2.png}
% \end{center}
% \end{minipage}
%  \caption{The painted parts correspond to the left side of (\ref{Sommerfeld}). Left figure shows the size surrounded by closed trajectory, right figure shows the size surrounded by one periodic opened trajectory and horizontal axis.}
%  \label{trajectory}
%\end{figure}

Before ending this section, let us note the periodic conditions
of the coordinates and the quantization condition.
Since the two original variables, $(\chi_1,\chi_2)$,
are independent periodic variables of the period $2\pi$,
in addition to the trivial periodicity of $2\pi$ for $\Phi$ and of $4\pi$
for $\phi$,
we have periodic conditions for
$(\Phi,\phi)\rightarrow (\Phi\pm \pi,\phi\pm 2\pi)$ and
$(\Phi,\phi)\rightarrow (\Phi\pm \pi,\phi\mp 2\pi)$.
The former (latter) corresponds to $\chi_2\pm 2\pi$ ($\chi_1\pm 2\pi$)
with $\chi_1$ ($\chi_2$) being fixed.
For open TDHFB trajectory (e.g., Fig.~\ref{fig:phase_space}),
the quantization condition becomes
\begin{equation}
 \mathcal{T}_{N,E_k}(\Phi,T/2)
	= \pm \frac{N}{2} \pi + \int_{-\pi}^\pi j d\phi
	= 2k\pi .
\end{equation}
This leads to the following:
\begin{equation}
	\int_{-\pi}^\pi j d\phi = 
	\begin{cases}
		2k\pi & \mbox{ for } N=4n ,\\
		(2k+1)\pi & \mbox{ for } N=4n+2 ,
	\end{cases}
	\label{EBK_2}
\end{equation}
where $k$ and $n>0$ are integer numbers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Canonical quantization}
\label{sec:canonical}

The most common approach to the quantization of the nuclear collective model
is the canonical quantization \cite{BM75}.
In the pairing collective model, the canonical quantization
was adopted in previous studies \cite{BBPK70,delta1,delta3}.
Assuming magnitude and phase of the pairing gap as collective coordinates,
a collective Hamiltonian was constructed in the second order in momenta.
Then, the Hamiltonian was quantized by the canonical quantization
with the Pauli's prescription.
In this section, we apply similar quantization method to
the TDHFB Hamiltonian (\ref{TDHFB_Hamiltonian_3}).
The main difference is that the collective canonical variables are
not assumed in the present case, but are obtained from the TDHFB dynamics
itself.
%The most natural procedure for quantizing classical dynamics is canonical quantization. The canonical variables become to operators and fulfill the commutation relation

The quantization method is not unique.
A common choice in the nuclear collective model is the Pauli's prescription.
However, it is not straightforward to apply it to the present case, 
because the TDHFB Hamiltonian (\ref{TDHFB_Hamiltonian_3})
is not limited to the second order in momenta.
In the present study, we adopt a simple symmetrized ordering, as
\begin{align}
	H(\hat{\phi},\hat{j},\hat{J})
	&= \sum_{l=1,2} \Omega_l \left\{ \epsilon_l (1-q_l) 
	- \frac{g}{4} (\Omega_l(1-q_l^2)+(1-q_l)^2) \right\}
	\nonumber \\
	&- \frac{g}{4}\Omega_{1}\Omega_{2}
	\left\{ \sqrt{(1-q_{1}^2)(1-q_{2}^2)}\cos{\hat{\phi}}
	+ \cos{\hat{\phi}}\sqrt{(1-q_{1}^2)(1-q_{2}^2)} \right\} .
\label{canonical_quantized_H}
\end{align}
As in Eq. (\ref{q_l}), $q_l$ contain $J$ and $j$ which
are replaced by
\begin{equation}
%  [\hat{\Phi},\hat{\Pi}] = i, \hspace{2em} [\hat{\phi},\hat{\pi}] = i .
	\hat{J} = -i\frac{\partial}{\partial\Phi},\quad\quad
	\hat{j} = -i\frac{\partial}{\partial\phi} .
\end{equation}
%Because quantized Hamiltonian must be Hermitian, 
%the ordering in mixing term of $\phi$ and $\pi$ is not unique. 
%However, we apply the quantization procedure of Hamiltonian in (\ref{2level}) as follows
Since $\Phi$ is a cyclic variable, 
we write the collective wave function $\Psi(\Phi,\phi)$ as eigenstates
of the particle number $N$ in a separable form
\begin{equation}
  \Psi_k^{(N)}(\Phi,\phi) = 
	\frac{1}{\sqrt{2\pi}}e^{i\frac{N}{2}\Phi}\psi_k^{(N)}(\phi) .
\end{equation}
Then, the problem is reduced to the one-dimensional Schr\"{o}dinger
equation for the motion in the relative angle $\phi$.
%pairing vibrational mode can be obtained by solving the Schrodinger equation
\begin{equation}
	H\left( \phi,-i\frac{d}{d\phi},\frac{N}{2} \right)
	\psi_k^{(N)}(\phi) = E_k^{(N)}\psi_k^{(N)}(\phi),
	\label{Schroedinger_eq}
\end{equation}
%where $i$ is the excitation index under the system of fixed total particle number $N$. To solve the Schrodinger equation, the wave function can be expanded as
%we consider two different methods, which are classified by Type I and Type II.\par
%Type I: periodic boundary condition.

The wave function should have a periodic property with respect to
the variable $\phi$;
$\psi_k(\phi)=\psi_k(\phi+4\pi)$.
For the adopted simple ordering of Eq. (\ref{canonical_quantized_H}),
it is convenient to use the eigenstates of $\hat{j}$ as the basis
to diagonalize the Hamiltonian.
They are
\begin{equation}
%  \hat{\pi}\chi_\nu(\phi) = \nu \chi_\nu(\phi) \\
	\chi_p(\phi) = \frac{1}{\sqrt{4\pi}} e^{ip\phi} ,
	\quad\quad\mbox{with $p$: integer or half integer} .
\end{equation}
Since the Hamiltonian (\ref{canonical_quantized_H}) contains only
terms linearly proportional to $e^{\pm i\phi}$,
the basis states $\chi_p$ with half-integer difference in $p$
are not coupled with each other.
Thus, the eigenstates of Eq. (\ref{Schroedinger_eq}) can be expanded as
\begin{equation}
	\psi_k^{(N)}(\phi) = 
	\sum_{p=p_{\rm min},p_{\rm min}+1,\cdots}^{p_{\rm max}}
	c_{k,p}^{(N)} \chi_p^{(N)}(\phi) .
	\label{CQ_eigenstate}
\end{equation}
According to the relation $j=(n_2-n_1)/4=(N-2n_1)/4$ in Eq. (\ref{momentum}),
we adopt the (half-)integer values of $p$ for $N=4n$ ($N=4n+2$)
with integer $n$.
This is consistent with the quantization condition (\ref{EBK_2}).
The coupling term with different $p$
in Eq. (\ref{canonical_quantized_H}) vanishes
for $n_l=0$ and $n_l=2\Omega_l$,
which restricts values of $p$ in a finite range of
$p_{\rm min}\leq p \leq p_{\rm max}$.

%where $\nu_0$ is the minimum value of $\nu$.
%$\chi_m(\phi)$ is the eigenfunction of momentum operator fulfills
%\begin{align}
%  \hat{\pi}\chi_m(\phi) = p_m \chi_m(\phi) \\
%  \chi_m(\phi) \propto e^{ip_m\phi} .
%\end{align}
%Due to the periodic boundary condition in coordinate space of Hamiltonian
%\begin{equation}
%  H(\phi) = H(\phi+2\pi)
%\end{equation}
%and the integer or half-integer number of $p_m$ in maximum and minimum value, the wave function fulfills
%\begin{equation}
%  \psi_i(\phi) = 
%  \left\{ 
%  \begin{array}{c}
% \psi_i(\phi+2\pi) \hspace{2em} ({\rm for} \hspace{1em} p_m=m) \\    
% -\psi_i(\phi+2\pi) \hspace{2em} ({\rm for} \hspace{1em}  p_m=m/2) \\    
%  \end{array}
%  \right.  
%\end{equation}
%and $p_m$ must increase or decrease in increments of one 
%\begin{align}
%  p_{min}, p_{min}+1, \cdots ,p_{max}-1, p_{max} .
%\end{align}
%Therefore, we can easily diagonalize the Hamiltonian in the basis of momentum eigenfuction.\par
\begin{comment}
Type II: adiabatic solution. We employ a canonical transformation
\begin{align}
  Q = Q(\phi,\pi) = \pi, \hspace{2em} P = P(\phi,\pi) = -\phi
\end{align}
assuming $\pi$ as a coordinate and $\phi$ as a conjugate momentum. Because wave function is expected to be localized in $\phi$ space when pairing correlation is strong, we expand $\phi$ up to second order. The change in Hamiltonian is only
\begin{align}
  \cos{\phi} = \cos{P} \to 1 + \frac{1}{2}\frac{\partial^2}{\partial Q^2} .
\end{align}
In this solution, there is no periodic boundary condition, which opposite to Type I. \par
\end{comment}

In order to estimate the two-particle transfer matrix elements,
we construct the corresponding operators as follows.
The classical form of matrix elements are obtained as
\begin{align}
  S_+(\Phi,J;\phi,j) &= \braket{Z|\hat{S}_+|Z} = \left( S_1\sqrt{1-q_1^2}e^{-i\phi/2}+S_2\sqrt{1-q_2^2}e^{i\phi/2} \right) e^{i\Phi}
	\label{Sp_mean_value} \\
  S_-(\Phi,J;\phi,j) &= \braket{Z|\hat{S}_-|Z} = \left( S_1\sqrt{1-q_1^2}e^{i\phi/2}+S_2\sqrt{1-q_2^2}e^{-i\phi/2} \right) e^{-i\Phi} .
	\label{Sm_mean_value}
\end{align}
%For simplicity, we ignore the part of off-diagonal element (e.g. $\braket{Z'|\hat{S}_+|Z}$). 
%To requantize the two-particle transition operators, 
Again, we adopt a simple symmetrized ordering for the quantization:
\begin{eqnarray}
%  &S_+(\Phi,\Pi;\phi,\pi) \to S_+(\hat{\Phi},\hat{\Pi};\hat{\phi},\hat{\pi}) \nonumber \\
  S_\pm (\hat{\Phi},\hat{J};\hat{\phi},\hat{j})
	&=& \frac{1}{4}\left( \Omega_1\sqrt{1-q_1^2}e^{\mp i\hat{\phi}/2}
+\Omega_2\sqrt{1-q_2^2}e^{\pm i\hat{\phi}/2} \right) e^{\pm i\hat{\Phi}} 
	\nonumber \\
&+ & \frac{1}{4}e^{\pm i\hat{\Phi}}\left( 
e^{\mp i\hat{\phi}/2}\Omega_1\sqrt{1-q_1^2}
+e^{\pm i\hat{\phi}/2}\Omega_2\sqrt{1-q_2^2} \right) .
\end{eqnarray}
The exponential factors $e^{\pm i\Phi}$ change
the total particle number $N\rightarrow N\pm 2$,
while $e^{\pm i\phi/2}$ change the relative numbers,
$n_2-n_1 \rightarrow n_2-n_1 \pm 2$.
%For example, the pair-additional transition strength in requantized form is
Using these operators,
the pair-addition transition strengths are calculated as
\begin{align}
B(P_{\rm ad};k\rightarrow k') &=
|\braket{N',k'|S_+(\hat{\Phi},\hat{J};\hat{\phi},\hat{j})|N,k}|^2
	\nonumber \\
&= \left| \int_{0}^{2\pi}d\Phi \int_{-2\pi}^{2\pi}d\phi 
	\psi^{(N')*}_{k'}(\phi) e^{-i\frac{N'}{2}\Phi}
	S_+(\hat{\Phi},\hat{J};\hat{\phi},\hat{j})
	\psi^{(N)}_{k}(\phi)e^{i\frac{N}{2}\Phi} \right|^2,
\end{align}
which automatically vanishes for $N'\neq N+2$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier decomposition of time-dependent matrix elements}
\label{sec:Fourier}

The requantization and calculation of the matrix elements can be also
performed using the time-dependent solutions of the TDHFB.
It was proposed and applied to the two-level pairing model \cite{CDS84},
which we recapitulate in this section.

The TDHFB provides a time-dependent solution $Z(t)$ starting from
a given initial state $Z(0)$.
The energy eigenvalues and the corresponding closed trajectories
are determined from the EBK quantization condition (\ref{EBK}).
The pair transfer matrix elements are evaluated as the Fourier components
of the time-dependent mean values $S_\pm(t)=S_\pm(Z(t))$,
Eqs. (\ref{Sp_mean_value}) and (\ref{Sm_mean_value}).
%In the ground state, the motion in the global gauge angle $\Phi$
%is independent from the $\phi$ motion.
Since the global gauge angle $\Phi$ is a cyclic variable,
the motion in the relative gauge angle $\phi$ is independent from $\Phi$.
Thus, we calculate the time evolution of $\phi(t)$, and
find the period of the $k$-th closed trajectory $T$
satisfying Eq. (\ref{EBK}).
Then, the Fourier component,
\begin{equation}
	\tilde{S}_\pm(E_k; \omega) = \frac{1}{T}\int_0^T dt
	e^{i\omega t} S_\pm(t) ,
\label{Fourier_decomposition}
\end{equation}
corresponds to the pair transfer matrix element from the state $k$ to $k'$
when $\omega=2\pi (k'-k)/T$.
The pair-addition transition strengths are calculated as
\begin{equation}
	B(P_{\rm ad};k\rightarrow k') 
	= \left|\tilde{S}_+\left(E_k;\frac{2\pi}{T}\Delta k \right) \right|^2 ,
\end{equation}
with $\Delta k=k'-k$.
In this approach, the transition between the ground states of
neighboring nuclei ($N\rightarrow N+2$) corresponds to
the stationary component ($k=0$ and $\Delta k=0$),
namely the expectation value in the BCS approximation.

The derivation  of Eq. (\ref{Fourier_decomposition})
is based on the wave packet in the
classical limit \cite{LL65}.
The TDHFB state is assumed to be a superposition of eigenstates
in a narrow range of energy $E_{k_0}-\Delta E<E_k<E_{k_0}+\Delta E$,
\begin{equation}
	\ket{Z(t)} = \sum_N \sum_k c_k^N \ket{\phi_k^N} e^{-iE_k t} ,
	\label{superposition}
\end{equation}
where the eigenenergies are evenly spaced and
the coefficients $c_k^N$ slowly vary with respect to $k$ and $N$.
The expectation value of $S_\pm$ is
\begin{equation}
	S_\pm(t) = \sum_{N} \sum_{k,k'} c_{k'}^{N+2*} c_k^N 
	\braket{\phi_{k'}^{N+2}|S_\pm|\phi_k^N} e^{i(E_{k'}-E_k)t} .
\end{equation}
The matrix element $\braket{\phi_{k'}^{N+2}|S_\pm|\phi_k^N}$
quickly disappears as $|k'-k|$ increases, while it stays almost constant
for the small change of $k$ and $N$ with $|k'-k|$ being fixed.
Thus, we may approximate $c_{k'}^{N+2}\approx c_k^N$,
$E_{k'}-E_k\approx \omega_0 \Delta k$, and
that $\braket{\phi_{k'}^{N+2}|S_\pm|\phi_k^N}\approx
\braket{\phi_{k_0+\Delta k}^{N+2}|S_\pm|\phi_{k_0}^N}$ 
\begin{equation}
	S_\pm(t) \approx \sum_{N} \sum_{k} \left|c_k^N\right|^2
	\sum_{\Delta k}
	\braket{\phi_{k_0+\Delta k}^{N+2}|S_\pm|\phi_{k_0}^N}
	e^{i\omega_0 \Delta k t} 
	= \sum_{\Delta k}
	\braket{\phi_{k_0+\Delta k}^{N+2}|S_\pm|\phi_{k_0}^N}
	e^{i\omega_0 \Delta k t}  ,
\end{equation}
where $k_0$ is a representative index value of the superposition in 
Eq. (\ref{superposition}).
From this classical wave packet approximation,
we obtain Eq. (\ref{Fourier_decomposition}).
It is not trivial to justify the approximation for small values of $\Omega$
and for transitions around the ground state.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}

In this section, we study the seniority-zero states ($\nu_1=\nu_2=0$) in
the two-level system with equal degeneracy, $\Omega_1=\Omega_2=\Omega$.
Since all the properties are scaled with the ratio, $g/\Delta\epsilon$,
where $\Delta\epsilon$ is the level spacing
$\Delta\epsilon=\epsilon_2-\epsilon_1$,
we define a dimensionless parameter to control the strength of
the pairing correlation;
\begin{align}
  x = 2\Omega\frac{g}{\Delta\epsilon} .
\end{align}
For sub-shell-closed systems with $N=2\Omega$ system,
the transition from normal ($x<1$) to superfluid ($x>1$) states
takes place at $x=1$

%We consider two types of system: [1. Ideal type] $\Omega$ is large, which causes BCS approximation works very well for $x>>1$; [2. Realistic type] $\Omega$ is comparable to the capacity of single particle level between neighborhood magic numbers in nuclei. For each type, we consider the dynamics for $N=\Omega$ (half-filled) and $N=2\Omega$ (closed shell) system.\par

%In excitation energy spectra, we also used harmonic approximation to compare with both requantization methods. The excitation energy in harmonic approximation is expressed by mass parameter and second derivative of potential
%\begin{align}
%  \omega = \sqrt{\frac{V''(\pi)}{B(\pi)}}
%\end{align}
%where
%\begin{align}
%\left\{
%\begin{array}{c}
%  V(\pi) = \sum_{l=1,2} \epsilon_l(-2S_lq_l+\Omega_l) - g\sum_{l=1,2} \frac{S_l}{2}(2S_l(1-q_l^2)+(1-q_l)^2) - 2gS_{1}S_{2}\sqrt{(1-q_{1}^2)(1-q_{2}^2)} \\
%  B^{-1}(\pi) = 2gS_{1}S_{2}\sqrt{(1-q_{1}^2)(1-q_{2}^2)}
%\end{array}
%\right. .
%\end{align}
%Because of one-dimensional system, we only obtain the lowest excitation energy from harmonic approximation.

%The two-nucleon transfer reaction is a specific probe of pairing correlation. We define pair additional transition strength as
%\begin{align}
%  B(Pad) \equiv |\braket{N+2,\beta|S_+|N,\alpha}|^2
%  \label{Pad}
%\end{align}
%where $\alpha,\beta$ are the excitation index in fixed particle number system. 
%\begin{comment}
%To estimate the requantization method, we also define the ratio of transition strength
%\begin{align}
%  r \equiv \frac{B(Pad)_{req}}{B(Pad)_{exact}}.
%  \label{ratio}
%\end{align}
%$B(Pad)_{req}$ indicates the result from canonical quantization or semiclassical approach. The requantization result is perfect when $r=1$.
%\end{comment}

We apply the requantization methods in Sec.~\ref{sec:requantization}.
In the following, the stationary-phase approximation to the path
integral in Sec.~\ref{sec:SPA} is denoted as ``SPA'',
the Fourier decomposition method (Sec.~\ref{sec:Fourier}) as ``FD'',
and the canonical quantization (Sec.~\ref{sec:canonical}) as ``CQ''.
In addition, the harmonic approximation at the ground state is
called ''HA''.
Note that the SPA and the FD produce the same eigenenergies which
are based on the EBK quantization rule.

\subsection{Large-$\Omega$ cases}

In the limit of $\Omega\rightarrow\infty$,
we expect that the classical approximation becomes exact.
Here, we adopt $\Omega=50$ with $N=100$ (closed-shell configuration)
and $N=50$ (mid-shell configuration).

%The system consists of $\Omega=50$ and $\Delta\epsilon=1$(MeV). We show the results in $N=50$ system, which is always superfluidity in ground state.\par
Calulated excitation energies are shown in Figs.~\ref{fig:N50energy}
and \ref{fig:N100energy}.
The results of SPA/FD and CQ are compared with the exact values.
At the weak pairing limit of $x\rightarrow 0$,
the excitation energies are multiples of $2\Delta\epsilon$, 
which correspond to pure $2n$-particle-$2n$-hole excitations.
Both the weak and the strong pairing limits
are nicely reproduced by all the calculations,
while the CQ method produces excitation energies slightly lower than the
exact values in an intermediate region around $x=1$.
It is somewhat surprising to see that the deviation is larger for
the case of the mid-shell configuration ($N=50$) than the closed shell
($N=100$).

%Both requantization methods converge to exact solution in no pairing and strong pairing limit. However, we can see the result from canonical quantization is under estimated around $x=1$, while semiclassical approach is almost perfect in all region. The lowest excitation energy obtained from harmonic approximation well reproduces exact result.

The deviation in the CQ method is mainly due to the zero-point energy
in the ground state.
Since we solve the collective Schr\"odinger equation (\ref{Schroedinger_eq})
with the quantized Hamiltonian of Eq. (\ref{canonical_quantized_H}),
the zero-point energy $\Delta E>0$ is inevitable in the CQ method.
The $\Delta E$ is associated with the degree of localization of
the wave function.
Thus, the magnitude of $\Delta E$ for ``bound'' states are different
from that for ``unbound'' states.
See Fig.~\ref{fig:phase_space}.
In the strong pairing limit, the potential minimum is deep enough to bound
both ground and excited states.
Conversely, all the states are unbound in the weak limit.
In both limits, $\Delta E$ for ground and excited states are similar,
and they are canceled for the excitation energy.
However, near $x=1$, the ground state is bound,
while the excited states are unbound.
In this case, $\Delta E$ is larger in the ground state than in the
excited states, which makes the excitation energy smaller.
This also explains the difference between the mid-shell and closed-shell
configurations.
In the closed shell, all the states are unbound for $x<1$,
while, in the mid-shell,
there is a region in $x<1$ where the ground state is bound but
the excited state is unbound.

The obtained wave functions in the SPA and the CQ can be decomposed in the
$2n$-particle-$2n$-hole components in Fig.~\ref{fig:N50_occ}.
In the SPA, it is done as Eq. (\ref{semi_wave_func1}) and
the normalized squared coefficients $|C_m^{(E_k,J)}|^2$ are plotted
in Fig~\ref{fig:N50_occ}.
For the CQ, $|c_{k,p}^{(N)}|^2$ in Eq. (\ref{CQ_eigenstate}) are shown.
Here, $m$ and $p$ are related to each other, $2p=J-2m$.
They show excellent agreement with the exact results, not only for the
ground state but for excited states.
We find the SPA is even more precise than the CQ.

Next, let us discuss the transition matrix elements.
The calculated pair-addition strengths $B(P_{\rm ad})$ are shown in
Fig.~\ref{fig:N50Pad} for $N=98\rightarrow 100$ and $N=48\rightarrow 50$.
In this paper, we discuss only $k=0$ (ground state) and $k=1$
(1st excited $\nu=0$ state).
The FD calculation is based on the time evolution of the expectation 
value $S_+(t)$ with fixed $(J,\Phi)$ in Eq. (\ref{Fourier_decomposition}).
%There is some ambiguity for the calculation of $S_+(t)$,
%namely, for which trajectory we should calculate $S_+(t)$.
For $(N,k)\rightarrow (N+2,k')$ transitions,
we basically adopt the trajectories for the initial state, namely,
the one with $J=N/2$ satisfying the $k$-th EBK quantization condition.
The $k\rightarrow k$ ($\Delta k = 0$)  transitions
correspond to the intraband transitions of the
pair-rotational band, when the state is deformed in the
gauge space (pair deformation).
For the ground-state band ($k=0$),
this is nothing but the expectation value at the BCS wave function.
The calculation trivially leads to vanishing interband transitions for 
$k=0\rightarrow k'$ ($k'\neq 0$), because the $S_+(t)$ is stationary
in the ground state.
Thus, for the interband transition of $(N,k=0)\rightarrow (N+2,k=1)$
transitions, the trajectory satisfying the EBK condition of $k=1$ is used to
perform the Fourier decomposition (\ref{Fourier_decomposition})
of $\omega=2\pi/T$.

Near the closed-shell configuration ($N=98\rightarrow 100$),
the pair-addition strengths for the intraband transitions ($\Delta k=0$)
drastically increase around $x=1$.
This reflects a character change from the pair vibration ($x\lesssim 1$)
to the pair rotation ($x\gtrsim 1$).
The $B(P_{\rm ad}; k\rightarrow k)$ in the pair-rotational transitions
are about 20 times larger than those in the vibrational transitions.
The interband $B(P_{\rm ad}; 0\rightarrow 1)$ are similar to
the $B(P_{\rm ad}; 0\rightarrow 0)$ in the vibrational region
($x\lesssim 1$), because they both change the number of pair-phonon quanta
by one unit.
In contrast, $B(P_{\rm ad}; 1\rightarrow 0)$, which change the phonon
quanta by two, are almost zero.
In the pair-rotational region ($x \gg 1$),
$B(P_{\rm ad}; 1\rightarrow 0)$ and $B(P_{\rm ad}; 0\rightarrow 1)$
are roughly identical,
while their magnitude is significantly smaller than
$B(P_{\rm ad}; 0\rightarrow 1)$ at $x\lesssim 1$.
This indicates the character change of the $k=1$ states
from the vibration around the ``spherical'' ground state
to that around the ``deformed'' ground state.
This is analogous to the quadrupole vibrations in spherical nuclei
and the beta vibrations in deformed nuclei, respectively.
The latter may correspond to the Higgs amplitude mode.

In the mid-shell region ($N=48\rightarrow 50$),
the intraband $B(P_{\rm ad}; k\rightarrow k)$ are smoothly increase as
$x$ increases.
Their values are larger than the interband strengths by about one order
of magnitude, indicating the pair-rotational character.
The interband $B(P_{\rm ad}; 0\rightarrow 1)$ shows a gradual decrease
as a function of $x$, while
$B(P_{\rm ad}; 1\rightarrow 0)$ are negligibly small,
even at $x\gg 1$.
This present a prominent difference from the closed-shell case.

All the features are nicely reproduced in the SPA method,
for both the closed- and mid-shell configurations.
The CQ method qualitatively agree with the exact calculation.
For instance, the order-of-magnitude difference between intraband
and interband transitions.
However, the precision of the CQ method is not so good,
especiallyaround $x=1$.
The FD method properly describes the main features
in the superflud phase, while it fails for the normal phase
($x\lesssim 1$). 
In the mid-shell configuration, the ground state is always
in the superfluid phase at $x>0$,
while the $k=1$ excited state corresponds to
the open (closed) trajectory at $0<x\lesssim 1$ ($x\gtrsim 1$).
For the open trajectory, the FD produces wrong values.
However, somewhat surprisingly,
the SPA results, which use these open trajectories for the construction
of wave functions,
reproduce main features of the exact results.

%The two-particle transition strength from $N=48$ to $N=50$ is shown in Figure \ref{fig:N50Pad}. We plotted four types of transition, $\ket{0_1^+}\to\ket{0_1^+}$, $\ket{0_1^+}\to\ket{0_2^+}$, $\ket{0_2^+}\to\ket{0_1^+}$ and $\ket{0_2^+}\to\ket{0_2^+}$. $\ket{0_1^+}\to\ket{0_1^+}$ and $\ket{0_2^+}\to\ket{0_2^+}$ correspond to pairing rotational transition, while $\ket{0_1^+}\to\ket{0_2^+}$ and $\ket{0_2^+}\to\ket{0_1^+}$ correspond to pairing vibrational transition. We can find the pairing rotational transition are much stronger than pairing vibrational transition in whole region. The minuscule strength of $\ket{0_2^+}\to\ket{0_1^+}$ is caused from peculiarity of two-level system. In canonical quantization, the pairing rotational transition is quantitatively reproduced, while the pairing vibrational transition is under estimated in $x<1.5$. In semiclassical approach, not only the pairing rotational transition is almost perfectly reproduced, but also the pairing rotational transition has the same tendency with exact solution. Because $\ket{0_2^+}$ has one node, in $\ket{0_1^+}\to\ket{0_2^+}$, most of component of matrix element are canceled with each other, causing weak transition strength. Hence the under estimation of $\ket{0_1^+}\to\ket{0_2^+}$ in semiclassical approach is not serious.\par
%From numerical results in ideal case, even though semiclassical approach is better than canonical quantization, canonical quantization is still able to describe physical values quantitatively. This is due to high collectivity of pairing in large $\Omega$ system.

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=60mm,angle=-90]{images/N50ex_energy.eps}
 \end{center}
 \caption{Excitation energies of $\ket{0_2^+}$, $\ket{0_3^+}$ and $\ket{0_4^+}$ in $N=50$ system.}
 \label{fig:N50energy}
\end{figure}

\begin{figure}[htbp]
 \begin{minipage}{1\hsize}
 \begin{center}
%  \includegraphics[width=67mm,angle=-90]{images/N50Xeq0p5occ_woType2.eps}
 \end{center}
 \end{minipage}
 \begin{minipage}{1\hsize}
 \begin{center}
%  \includegraphics[width=67mm,angle=-90]{images/N50Xeq2occ_woType2.eps}
 \end{center}
 \end{minipage}
 \caption{Occupation probability in excited $0^+$ states as a function of $\pi$ in $N=50$ system. The upper and lower panels display the results for $x=0.05$ and $x=0.2$, respectively. From the physical meaning of $\pi$, $-12.5$ corresponds to 0p50h configuration, $-11.5$ corresponds to 2p52h configuration, and so on.}
 \label{fig:N50_occ}
\end{figure}

\begin{figure}[htbp]
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N50Pad_exact.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N50Pad_Type1.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N50Pad_semiclassical.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \caption{The strength of pair additional transition from $N=48$ to $N=50$ system defined in (\ref{Pad}). The results of $\ket{0_1^+}\to\ket{0_1^+}$, $\ket{0_1^+}\to\ket{0_2^+}$, $\ket{0_2^+}\to\ket{0_1^+}$ and $\ket{0_2^+}\to\ket{0_2^+}$ are shown. Left panels: exact solution; Middle panels: canonical quantization; Right panels: semiclassical approach.}
 \label{fig:N50Pad}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsection{Small-$\Omega$ cases}

Next, we discuss systems with smaller degeneracy $\Omega=8$.
Again, we study systems near the closed-shell and the mid-shell configurations.
%$N=8$ system, which is always superfluidity in ground state; (2) $N=16$ system, which has phase transition at $x=1$ in ground state.\par

\subsubsection{Mid-shell configuration}

The calculated excitation energies are shown in Fig.~\ref{fig:N8energy}
for the $N=8$ case.
The SPA/FD reproduces the exact calculation in the entire region of $x$,
not only for the lowest but also for higher excited states.
The CQ reproduces the exact result in a weak pairing region ($x\lesssim 1$),
while it underestimates the excitation energies at $x\gtrsim 1$.
Analogous to the case of $\Omega=50$,
this is mainly due to effect of the zero-point energy $\Delta E$.
The ground-state energy in the CQ calculation 
is bound at $x\gtrsim 1$ ({\bf need to check}).
However, because of the weak collectivity with $N=8$,
the first excited state stays unbound even at the maximum $x$ in 
Fig.~\ref{fig:N8energy}.
Therefore, the energy shift $\Delta E>0$ is larger in the ground state,
which makes the excitation energy smaller.

The wave functions are plotted in Fig.~\ref{fig:N8_occ}. 
At the weak pairing case of $x=0.5$,
both the SPA and the CQ reproduce the exact result.
At $x=2$, the squared coefficients of the ground state is peaked at the
lowest $p$, which suggests that the state is not deeply bound in the
potential.
It is in contrast to the symmetric shape in Fig.~\ref{fig:N50_occ}.
The wave functions obtained by the CQ method has noticeable deviation
from the exact result.
On the other hand, the SPA wave functions are almost identical to the
exact ones.

%At no pairing limit, the excitation energies are multiples of $2\Delta\epsilon$. Both requantization methods converge to exact solution in no pairing limit. However, we can see the result from canonical quantization is under estimated after $x=1$, while semiclassical approach is almost perfect in all region. The lowest excitation energy obtained from harmonic approximation well reproduces exact result. 
%We can find both requantization methods reproduce well in $x=0.5$, and semiclassical approach is better than canonical quantization in $x=2$.

The pair-addition transition strengths from $N=6$ to $N=8$ are
shown in Fig.~\ref{fig:N8Pad}.
The intraband $k\rightarrow k$ transitions increase and 
the interband $k\rightarrow k\pm1$ transitions decrease as
functions of $x$.
Their relative difference becomes
more than one order of magnitude at $x\gtrsim 2$.
Thus, even at relatively small $\Omega$ and $N$,
the intraband transitions in the pair rotation is qualitatively different
from the interband transitions.

We find the excellent agreement between the SPA and the exact calculations.
The first excited state corresponds to the open trajectory which
turns out to almost perfectly reproduces the exact wave function.
In contrast, this open trajectory produces results far from the exact
result in the FD method.
It produces almost vanishing the intraband
$B(P_{\rm ad}; 1\rightarrow 1)$ and the interband transitions.
The $B(P_{\rm ad}; 0\rightarrow 0)$ shows a qualitative agreement for
its behavior, but is significantly underestimated.
The CQ method also underestimates the intraband transitions.

%The minuscule strength of $\ket{0_2^+}\to\ket{0_1^+}$ is caused from peculiarity of two-level system. In canonical quantization, all transition strengths are failed to reproduce quantitatively. In semiclassical approach, however, both pairing rotational transition and pairing vibrational transition are reproduced quantitatively in all region. \par
%In above case, semiclassical approach is much better than canonical quantization. The low collectivity of pairing in small $\Omega$ system influences significantly the methods of requantization. Nevertheless, semiclassical approach can still describe physical values quantitatively as in ideal case.

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=60mm,angle=-90]{images/N8ex_energy.eps}
 \end{center}
 \caption{Excitation energies of $\ket{0_2^+}$, $\ket{0_3^+}$ and $\ket{0_4^+}$ in $N=8$ system.}
 \label{fig:N8energy}
\end{figure}

\begin{figure}[htbp]
 \begin{minipage}{1\hsize}
 \begin{center}
%  \includegraphics[width=67mm,angle=-90]{images/N8Xeq0p5occ_woType2.eps}
 \end{center}
 \end{minipage}
 \begin{minipage}{1\hsize}
 \begin{center}
%  \includegraphics[width=67mm,angle=-90]{images/N8Xeq2occ_woType2.eps}
 \end{center}
 \end{minipage}
 \caption{Occupation probability in excited $0^+$ states as a function of $\pi$ in $N=8$ system. The upper and lower panels display the results for $x=0.05$ and $x=0.2$, respectively. From the physical meaning of $\pi$, $-2$ corresponds to 0p8h configuration, $-1$ corresponds to 2p10h configuration, and so on.}
 \label{fig:N8_occ}
\end{figure}

\begin{figure}[htbp]
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N8Pad_exact.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N8Pad_Type1.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N8Pad_semiclassical.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \caption{The strength of pair additional transition from $N=6$ to $N=8$ system defined in (\ref{Pad}). The results of $\ket{0_1^+}\to\ket{0_1^+}$, $\ket{0_1^+}\to\ket{0_2^+}$, $\ket{0_2^+}\to\ket{0_1^+}$ and $\ket{0_2^+}\to\ket{0_2^+}$ are shown. Left panels: exact solution; Middle panels: canonical quantization; Right panels: semiclassical approach.}
 \label{fig:N8Pad}
\end{figure}

\subsubsection{Closed-shell configuration}

In the closed shell with $N=16$,
the minimum-energy trajectory changes at $x=1$ from
$p=-4$ (normal phase) to
the BCS minimum $p>-4$ and $\phi=0$ (superfluid phase).
At the transitional point ($x=1$), the harmonic approximation is
known to collapse, namely to produce zero excitation energy.
In Fig.~\ref{fig:N16energy},
this collapsing is avoided in all the calculations (SPA/FD and CQ).
The behaviors of the lowest excitation agree with the exact calculations,
while the CQ method substantially underestimates those for higher states.
This is again due to the difference in the zero-point energy in
the ground and the excited states.
In the CQ calculation, the first excited state is bound at $x\gtrsim 1.?$,
but the second excited state is unbound for $x\lesssim 3.?$.

Near the transition point from the open to closed trajectories,
the wave functions calculated with the SPA and CQ methods somewhat differ
from the exact ones.
In Fig.~\ref{fig:N16_occ}, the wave functions at $x=0.5$ and 2 are presented.
They agree with exact calculation at $x=0.5$.
In contrast, we find some deviations for the first excited state ($k=1$)
at $x=2$.
This is because the $k=1$ trajectory corresponding to the first excited state
changes its character from open to closed at $x\approx 1.8$ {\bf need to check}.
Therefore, the first excited wave function is difficult to reproduce in
the SPA, although the wave functions for
the ground and higher excited states show reasonable agreement.
The similar disagreement is observed for the ground state near $x=1$.

%The excitation energy is shown in Figure \ref{fig:N16energy}. At no pairing limit, the excitation energies are multiples of $2\Delta\epsilon$. Both requantization methods converge to exact solution in no pairing limit, while under estimate the exact solution after $x=1$. However, semiclassical approach is close to exact solution. The lowest excitation energy obtained from harmonic approximation emerges after $x>1$, and well reproduces exact result in strong pairing limit. 
%The occupation probability in such excited states is plotted in Figure \ref{fig:N16_occ}. We can find both requantization methods reproduce well in $x=0.5$, and both methods have some disagreement of $\ket{0_2^+}$ and $\ket{0_3^+}$ in $x=2$. 
%This disagreement is caused by large fluctuation close to phase transition point (in Figure \ref{phase_space}, the changing from closed trajectory to opened trajectory). After the phase transition, the occupation probability of $\ket{0_4^+}$ is also well reproduced in semiclassical approach.

Singular behaviors near the transition points can be also observed in
the pair-addition transition strengths ($N=14\rightarrow 16$) shown
in Fig.~\ref{fig:N16Pad}. 
At $x=1$, the intraband $B(P_{\rm ad}; 0\rightarrow 0)$ shows a kink
in the SPA,
and $B(P_{\rm ad}; 1\rightarrow 1)$ shows another kink at $x\approx 1.8$.
These exactly correspond to the transition points from open to closed
trajectories.
Nevertheless, the overall behaviors are well reproduced and
the values at the weak and strong pairing limit are well reproduced
in the SPA.
The CQ calculation also shows smoothed kink-like behaviors near the
transition points.
The FD method does not have a kink for $B(P_{\rm ad}; 0\rightarrow 0)$,
because $S_+(t)$ is calculated for a $N=14$ system.
Both intraband and interband transitions reasonably agree with
the exact results at $x\gtrsim 1.8$,
while the $k=1$ state is not properly reproduced at $x\lesssim 1.8$
with the open trajectory.

%Between pairing rotational transition and pairing vibrational transition, we can find the transition strength are compareable in $x<1$, and pairing rotational transition are dominant in $x>1$. The minuscule strength of $\ket{0_2^+}\to\ket{0_1^+}$ is caused from peculiarity of two-level system.  In canonical quantization, even though the tendency of $\ket{0_1^+}\to\ket{0_1^+}$ are reproduced, all transition strengths are failed to reproduce quantitatively. In semiclassical approach, the strengths have disagreement around phase transition region, but are recovering toward no pairing and strong pairing limit. \par
%In closed shell system, the phase transition is semiclassical approach is more quantitative than canonical approach except phase transition region.

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=60mm,angle=-90]{images/N16ex_energy.eps}
 \end{center}
 \caption{Excitation energies of $\ket{0_2^+}$, $\ket{0_3^+}$ and $\ket{0_4^+}$ in $N=16$ system.}
 \label{fig:N16energy}
\end{figure}

\begin{figure}[htbp]
 \begin{minipage}{1\hsize}
 \begin{center}
%  \includegraphics[width=67mm,angle=-90]{images/N16Xeq0p5occ_woType2.eps}
 \end{center}
 \end{minipage}
 \begin{minipage}{1\hsize}
 \begin{center}
%  \includegraphics[width=67mm,angle=-90]{images/N16Xeq2occ_woType2.eps}
 \end{center}
 \end{minipage}
 \caption{Occupation probability in excited $0^+$ states as a function of $\pi$ in $N=50$ system. The upper and lower panels display the results for $x=0.05$ and $x=0.2$, respectively. From the physical meaning of $\pi$, $-4$ corresponds to 0p0h configuration, $-3$ corresponds to 2p2h configuration, and so on.}
 \label{fig:N16_occ}
\end{figure}

\begin{figure}[htbp]
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N16Pad_exact.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N16Pad_Type1.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \begin{minipage}{0.33\hsize}
 \begin{center}
%  \includegraphics[width=70mm,angle=-90]{images/N16Pad_semiclassical.eps}
 \end{center}
 \captionsetup{labelformat=empty,labelsep=none}
 \end{minipage}
 \caption{The strength of pair additional transition from $N=14$ to $N=16$ system defined in (\ref{Pad}). The results of $\ket{0_1^+}\to\ket{0_1^+}$, $\ket{0_1^+}\to\ket{0_2^+}$, $\ket{0_2^+}\to\ket{0_1^+}$ and $\ket{0_2^+}\to\ket{0_2^+}$ are shown. Left panels: exact solution; Middle panels: canonical quantization; Right panels: semiclassical approach.}
 \label{fig:N16Pad}
\end{figure}
%In realistic type, semiclassical approach is much better than canonical quantization in every physical values, especially in two-particle transition. The most difficult case is closed shell system with BCS phase transition. However, semiclassical approach works well in closed shell system except phase transition region.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Collective model treatment}

In literatures, the collective model has been proposed
and utilized for the nuclear pairing dynamics.
For those studies, the pairing gap parameter is assumed to be the
collective coordinates.
This is analogous to the five-dimensional (5D) collective (Bohr) model,
in which the collective coordinates are assumed to be the
quadrupole deformation parameters, $\alpha_{2\mu}$.
The 5D collective model has been extensively studied and
successfully applied to analysis on numerous experimental data.
On the contrary, the collective model for the pairing was not so
successful in comparison with experimental data.
In this section, we examine the collective treatment of the pairing.

%The most popular collective coordinate to describe pairing vibration is pairing gap, which is the order parameter in U(1) gauge space. The algebra form of pairing vibration is analogous the vibration in real three-dimensional space which suppose to be the swinging in quadrupole deformation parameters $\beta,\gamma$. In two-level system, the pairing gap becomes
%\begin{align}
%\Delta &= g\braket{Z|\hat{S}_-|Z} \nonumber \\
%	&= g\sum_{l=1}^2 S_l\sqrt{1-q_l^2}e^{-i\chi_l} \nonumber \\
%	&= g \left( S_1\sqrt{1-q_1^2}e^{i\frac{\phi_1}{2}}+S_2\sqrt{1-q_2^2}e^{-i\frac{\phi_2}{2}} \right) e^{-i\Phi}.
%\end{align}
%. Without total gauge angle part $e^{i\Phi}$, $\Delta$ is still complex number in general. However, the pairing gap is real number with the definition by BCS wave function

In the superfluid phase, the energy minimization always lead to $\phi=0$.
Although the global gauge angle $\Phi$ is arbitrary,
the gap parameter is usually taken as real value ($\Phi=0$).
\begin{align}
\Delta(j,J) = g\braket{Z_0|\hat{S}_-|Z_0} %= g\sum_{k>0} u_kv_k \nonumber \\
	= g\sum_{l=1}^2 S_l\sqrt{1-q_l^2} . % = \Delta(\pi) .
 \label{Delta_BCS}
\end{align}
The relation between $j$ and $\Delta$ 
and that between $j$ and $E=\cal{H}(\phi=0,j;J)$
are shown for $\Omega=8$
in Figs. \ref{fig:p_Delta1} ($N=8$) and \ref{p_Delta2} ($N=16$).
%These results correspond to the system in Section 4.2. 
The gap $\Delta$ is largest at $j=0$ (equal filling in both levels),
and smallest at end points of $j$.
The constrained minimization with respect to $\Delta$ cannot 
produce the states corresponding to $j>0$.
Apparently, we cannot map the entire region of $j$ to $\Delta$,
since there is no one-to-one correspondence between $j$ and $\Delta$.

If the system has a deep potential at $j<0$ to prevent the wave function
from spreading toward the region of $j<0$,
we may justify the use of $\Delta$ as a coordinate in
the collective model.
For very large $\Omega$, this is the case.
However, for $\Omega=8$, the treatment can be justified only for
a very limited region of $x$, namely in the weak pairing limit.
For instance, we show the potential energy curves as functions of $\Delta$
in Fig.~\ref{fig:Delta_E} for the $N=16$ case.
In the weak pairing case ($x=0.5$),
assuming the Neumann boundary condition at $\Delta=0$,
we may obtain the colletive wave functions up to the third excited states.
On the other hand, in the strong pairing case ($x=3.2$), 
even the ground state may not be bound in the region of $j<0$.

%We also consider the behavior of energy surface.  After mapping, the energy surface at $\phi=0$
%\begin{align}
%  \left\{
%  \begin{array}{c}
%    E = E(\phi=0,\pi) \\
%    \Delta = \Delta(\pi)
%  \end{array}
%  \right. \hspace{1em} \to \hspace{1em} f(E,\Delta) = 0
%  \label{mapping}
%\end{align}
%can be connected with pairing gap $\Delta$. The implicit function form is necessary. On the other hand, we consider pairing gap as a collective coordinate ``by hand''. The collective coordinate can be derived from CBCS equation
%\begin{align}
% &\delta\braket{BCS|\hat{H}'|BCS} = 0, \hspace{1em} \hat{H}' = \hat{H} - \mu \hat{N} - \lambda \hat{S}_-
% \label{CBCS}
%\end{align}
%where $\lambda$ is Lagrange multiplier with respect to $\Delta$. In CBCS case, the energy surface is the explicit function of $\Delta$. The examples of energy surface from both mapping and CBCS are shown in Figure \ref{fig:Delta_E}. The results correspond to the system in Section 4.2, $N=16$ case. We found that both mapping and CBCS results are the same up to the turning point, which is infinite gradient point with respect to $\Delta$. However, CBCS cannot reproduce the upper surface after turning point because of no one-to-one correspondence between $\pi$ and $\Delta$. The pairing gap can only describe a part of the collective path. In two-level system, from \ref{Delta_BCS}, the gradient of $\Delta$ with respect to $\pi$ is 
%\begin{align}
% \frac{\partial\Delta}{\partial\pi} = \frac{q_2}{\sqrt{1-q_2^2}} - \frac{q_1}{\sqrt{1-q_1^2}} .
%\end{align}
%It represents, in $\pi$ space, the gradient goes to $+\infty$ at the left end and $-\infty$ at the right end. Hence the turning point $\frac{\partial\Delta}{\partial\pi}=0$ must exist on the way of $\pi$. Therefore, paining gap is not suitable collective coordinate to describe pairing vibration in two-level system.\par
%However, the effect of the turning point is not expected to be important in some cases. In the top panel of Figure \ref{fig:Delta_E}, we found that the low-lying energy spectra are located below the turning point. Because the energy of turning point is high, the collective path below the turning point is sufficient to describe the low-lying excited states. Therefore, the collective coordinate $\Delta$ is available to describe low-lying pairing vibration in weak pairing correlation system.

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=45mm,angle=-90]{images/N8p_Delta.eps}
 \end{center}
 \caption{Pairing gap $\Delta$ in (\ref{Delta_BCS}) as a function of $\pi$ with various pairing strength in $N=8$ system.}
 \label{fig:p_Delta1}
\end{figure}

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=45mm,angle=-90]{images/N16p_Delta.eps}
 \end{center}
 \caption{Pairing gap $\Delta$ in (\ref{Delta_BCS}) as a function of $\pi$ with various pairing strength in $N=16$ system.}
 \label{fig:p_Delta2}
\end{figure}

\begin{figure}[htbp]
 \begin{center}
%  \includegraphics[width=130mm,angle=-90]{images/Delta_compare.eps}
 \end{center}
 \caption{Energy surface with respect to $\Delta$. The results correspond to Section 4.2, $N=16$ system. Purple lines are obtained from (\ref{mapping}) and green dashed lines are obtained from (\ref{CBCS}). The blue lines are excitation energy spectra until 3rd excitation obtained from (\ref{Sommerfeld}). The top, middle and bottom panels display the energy surfaces with various pairing strength.}
 \label{fig:Delta_E}
\end{figure}

\newpage

\section{Conclusion}

The different methods of the requantization of the TDHFB dynamics
was studied for the two-level pairing model;
the stationary phase approximation (SPA) of the path integral,
the canonical quantization (CQ), and the Fourier decomposiion (FD)
of the time-dependent observables.
In this model, since the global gauge angle $\Phi$ is a cyclic variable,
the TDHFB dynamics can be described by the integrable classical dynamics.
After the pair-rotation variables $(\Phi,J)$ are separated,
the remaining degrees of freedom, $(\phi,j)$, describe
the pair-vibrational motion.

In systems with large degeneracy $\Omega$ and number of particles $N$,
all the quantization methods reasonably reproduce the results of the
exact calculation for excitation spectra.
It is more difficult to reproduce the two-particle transfer matrix elements.
Nevertheless, for the large $\Omega$ and $N$,
we obtain qualitative agreement with the exact results.
These are ideal cases,
but realistic situations may have smaller $\Omega$ and $N$
in the valence space.

In systems with relatively small $\Omega$ and $N$,
the agreement is less quantitative for the CQ and the FD,
especially for the two-particle transfer matrix elements.
In contrast, the SPA keeps its accuracy in the entire range of
pairing strengths.
One of the reasons of its sucess is due to the inclusion of the off-diagonal 
parts of the pair transfer operator,
by the explicit construction of the microscopic wave functions.
The CQ and FD calculate the pair transfer matrix elements using only
the diagonal part (expectation value) of the operator $\hat{S}_\pm$,
based on Eqs. (\ref{Sp_mean_value}) and (\ref{Sm_mean_value}).
This is a good approximation when the collectivity is so large that the
diagonal parts dominate.
However, the pairing collecitivity may not be too weak to justify this
treatment.

%To describe pairing collective motion by using collective coordinates, we applied requantization of TDHFB method. In classical picture of two-level system, we have two degrees of freedom. The total gauge angle describing pairing rotational mode is a cyclic variable in classical Hamiltonian. The intrinsic time dependent coordinate which is only one-dimensional  corresponds to pairing vibrational mode. 
%We discussed two requantization method, canonical quantization and semiclassical approach. In ideal type of system, both methods are reasonable even though semiclassical approach is better. In realistic type of system, the semiclassical approach is more quantitative than canonical quantization in excitation energy and wave function. 
%Especially, the transition strength of pair transfer has significant difference between two methods. 
%The off-diagonal part of pair transfer operator is important, causing requantization procedure lose a lot of information. 
%The advantage of semiclassical approach is that the wave function is microscopic, so that we can use original operator to calculate transition strength. Even though the system is as small as realistic case, the semiclassical approach works well except phase transition region.\par

We also investigated the conventional treatment of
the collective model which assumes that the collective coordinate is
the paring gap parameter.
As we mentioned before, the present two-level pairing Hamiltonian has
only one pair of collective variables $(\phi,j)$, in addition to the
pair rotational varialbles $(\Phi,J)$.
Even for such a simple system, we find that it is difficult to justify
the use of $\Delta$ as the collective coordinate,
especially for relatively small-$\Omega$ cases.
Basically, there is no one-to-one correspondence between $\Delta$ and $j$.
The collective wave functions are not necessarily bound in the region
where the variable $\Delta$ can represent.

% coordinate describing pairing vibration. We found that the collective path has significant difference compared with CBCS which use pairing gap as a constraint condition. The turning point must exist in energy surface, causing the collective path from CBCS only a part of the collective path in original coordinate. Hence pairing gap is not suitable collective coordinate to describe pairing collective motion in two-level system. However, if the energy of the turning point is much higher than low-lying excitation energy, pairing gap is available as a collective coordinate. In weak pairing correlation system, we expect that the low-lying excited states are able to be described by the coordinate of pairing gap. \par

Among the different requantization methods,
the SPA is the most accurate tool for desciption of
the pairing large amplitude collective motion in realistic nuclear system.
The weak point of this approach is that it is applicable only to the
integrable TDHFB system.
In order to solve this problem, we plan to first extract the
integrable collective submanifold in the many-dimensional TDHFB phase space.
For this purpose, the adiabatic self-consistent collective coordinate (ASCC)
method \cite{NMMY16} is a promising tool.
The combined study of the ASCC and the SPA for multi-level systems is our
next target under progress.

\bibliographystyle{unsrt}
\bibliography{ref}

\end{document}